Quasars Probing Quasars IX. The Kinematics of the Circumgalactic Medium Surrounding $z\sim2$ Quasars [v1.4]

### 2017-04-11 QPQ9/qpq9_draft.tex
contour plot now four levels

\begin{figure}
\includegraphics[width=3.5in]{Figures/fig_stacks_and_fits.pdf}
%\includegraphics[width=3.5in]{f02.pdf}
\caption{Mean and median absorption centered at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 of the foreground quasars for all QPQ9 pairs. The composites are shown in thin, 
black. Overplotted on the composites are Gaussian fits, normalized to pseudo-continua far away 
from a velocity of $0\,{\rm km\,s^{-1}}$ relative to $z_{\rm fg}$, and are shown in thick, blue. 
For the doublets, a second Gaussian with a fixed mean separation and a tied standard deviation is 
included in the modeling. The absorptions frequently exhibit large velocity widths. The dashed 
lines mark the centroids, which show small positive velocity offsets from $z_{\rm fg}$. The 
$1\sigma$ modeling error of the centroid and dispersion of the \ion{C}{2} mean stack is 
$\approx35\,{\rm km\,s^{-1}}$, which is also the typical modeling errors of the other stacks. 
}
\label{fig:stacks_and_fits}
\end{figure}

@@

In Figure~\ref{fig:stacks_and_fits}, we present mean and median stacks of \ion{C}{2}\,1334, 
\ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 absorption of the QPQ9 sample. We focus on the 
analysis results of the \ion{C}{2} mean stack. \ion{C}{4} and \ion{Mg}{2} are doublet transitions 
and it is more challenging to interpret their kinematics.
%To assess the average absorption of these transitions,
%we have combined the data with a novel technique.
%We have fitted the combined dataset using a cubic b-spline
%with knots at every 100\,kms.  We weighted each data
%point by the velocity width of the pixel to insure that 
%each sightline makes an equal contribution to the stack.
%These b-spline profiles are presented in the figure
%and a trivial integration gives the reported equivalent
%widths.
%[consider down-weighting systems with $<3\sigma$ EW]
%[Comnent on EW from 'standard' technique]
Two results are evident in Figure~\ref{fig:stacks_and_fits}: (i) the mean \ion{C}{2} stack 
exhibits excess absorption spanning a large velocity width; (ii) the mean absorption is likely 
skewed toward positive velocities. To model the absorption, we introduce a Gaussian profile while 
allowing a constant pseudo-continuum level to vary. We perform $\chi^2$ minimization with each 
channel given equal weight. From the best-fit to the data, we measure the centroid of the 
\ion{C}{2} stack to be $+208\,{\rm km\,s^{-1}}$ and the $1\sigma$ dispersion to be 
$312\,{\rm km\,s^{-1}}$. The centroid suggests an asymmetry that contradicts the standard 
expectation, while the dispersion suggests extreme kinematics. The median stack, on the other 
hand, shows weaker absorption, and the Gaussian model has a centroid with smaller offset and 
comparable dispersion.

We also create mean and median stack for the sub-sample with [\ion{O}{3}] redshifts and model the 
absorption with Gaussian best-fit. The \ion{C}{2} mean stack for this sub-sample has a centroid at 
$+239\,{\rm km\,s^{-1}}$, and a dispersion of $349\,{\rm km\,s^{-1}}$, consistent with the full 
sample.

@@

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_monte.pdf}
%\includegraphics[width=2.5in]{f05.pdf}
\caption{The same \ion{C}{2} mean stack shown in the first panel of 
Figure~\ref{fig:stacks_and_fits} is shown in thin, black. In thick, green, we overplot the 
Gaussian absorption model of the Monte Carlo simulations generated from a purely clustering 
argument. The model from clustering analysis is multiplied to the pseudo-continuum level of the 
stack of the observational data, shifted to the centroid of the stack of the data, and broadened 
by the mean redshift error in the data. The model is a good description of the data, with a 
dispersion within modeling error of the dispersion in the data. We overplot in dashed, yellow an 
absorption profile with $\sigma_v=654\,{\rm km\,s^{-1}}$, which is larger than the observed width 
by three times the standard deviation in the bootstrap analysis. Motions that produce a velocity 
width larger than this can be ruled out. We overplot in dashed, magenta an absorption profile with 
$\sigma_v=207\,{\rm km\,s^{-1}}$, which is smaller than the observed width by three times the 
modeling error. Unless gravitational and Hubble flows together with redshift error broadening 
produce a velocity width smaller than this, extra dynamical processes (e.g.\ outflows) will not be 
required to explain the observation. 
}
\label{fig:monte}
\end{figure}

Under the assumption that the intrinsic dispersion and the redshift uncertainty add in 
quadrature to give the observed width, we solve for the intrinsic dispersion in the \ion{C}{2} 
mean stack. For the full QPQ9 sample, with the mean 
$\sigma_{{\rm error}(z)}^{\rm full}=182\,{\rm km\,s^{-1}}$, we recover 
$\sigma_{\rm intrinsic}^{\rm full}=253\,{\rm km\,s^{-1}}$. For the sub-sample with [\ion{O}{3}] 
redshifts, we recover $\sigma_{\rm intrinsic}^{\rm [OIII]}=342\,{\rm km\,s^{-1}}$.

\cite{Eftekharzadeh+15} measured the clustering of quasars in the range $2.2<z<2.8$ while 
\cite{RodriguezTorres+16} measured the clustering of quasars in the range $1.8<z<2.2$. They 
estimated that these quasars are hosted by dark matter halos with mass $10^{12.5}\,M_\odot$ and 
$10^{12.6}\,M_\odot$ respectively. If dark matter halos hosting QPQ9 quasars have a characteristic 
mass $10^{12.6}\,M_\odot$ and follow an NFW profile \citep{NavarroFrenkWhite97} with concentration 
parameter $c=4$, at $z\approx2$ the maximum circular velocity is $345\,{\rm km\,s^{-1}}$. 
\cite{TormenBouchetWhite97} found that the maximum circular velocity is $\approx1.4$ times the 
maximum of the one-dimensional velocity dispersion $\sigma_{\rm 1D}$. Hence, the line-of-sight 
velocity dispersion typical of QPQ9 halos is $\sigma_{\rm 1D}=246\,{\rm km\,s^{-1}}$. We can 
investigate whether gravitational motions and Hubble flows are sufficient to reproduce the 
dispersion in the data, using Monte Carlo methods to simulate the absorption signals.

@@

In Figure~\ref{fig:monte}, we show a comparison of the observational data stack and the Gaussian 
absorption model of the Monte Carlo simulations. In the figure, the Gaussian absorption model is 
broadened by the mean redshift error by adding it in quadrature to the dispersion in the model. 
The resulting stack of mock spectra has a $1\sigma$ dispersion of $281\,{\rm km\,s^{-1}}$. This is 
within $1\sigma$ modeling error from the intrinsic dispersion in the \ion{C}{2} mean stack for the 
full sample, and about 2 times the modeling error away from the intrinsic dispersion in the stack 
of the sub-sample with [\ion{O}{3}] redshifts. We also test for the 
sensitivity of this measured dispersion to the correlation functions adopted. The QPQ6 clustering 
analysis is performed on only the strongest absorber near $z_{\rm fg}$, and a low $R_\perp$ 
sightline may in fact intercept more than one optically thick absorber. We double the number of 
absorbers for each mock sightline, and find the measured dispersion only increases by several 
${\rm km\,s^{-1}}$. Thus, the hypothesis that the observed velocity width is only produced by a 
combination of gravitational motions and Hubble flows cannot be ruled out. Extra dynamical 
processes are not necessary to explain the large velocity fields. From the bootstrap analysis of 
the observed dispersion, a width of $\sigma_v>654\,{\rm km\,s^{-1}}$ would be larger than the 
observed by three times the scatter (overplotted in Figure~\ref{fig:monte}). In other words, 
taking into account the redshift errors, motions in addition to gravitational and Hubble 
flows that will produce an intrinsic dispersion $\sigma_v>628\,{\rm km\,s^{-1}}$ can be ruled out.

On the contrary, $\sigma_{\rm 1D}<74\,{\rm km\,s^{-1}}$ together with Hubble velocities and 
broadening by redshift errors will result in a velocity width that is more than three times the 
modeling error away from the observed width (overplotted in Figure~\ref{fig:monte}). This implies 
that, unless the characteristic dark matter halo mass $<10^{11.0}\,M_\odot$, additional dynamical 
processes are not required to explain the observed width.

@@

\begin{figure}
\includegraphics[width=3.4in]{Figures/fig_stacks_fg.pdf}
%\includegraphics[width=3.4in]{f06.pdf}
\caption{Mean stacks of the foreground quasar spectra at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 for the QPQ9 sample. For \ion{C}{2} and \ion{Mg}{2}, the mean absorption is 
weaker than that in the background stacks, and there is no evidence for an excess at $z_{\rm fg}$. 
The stack for \ion{C}{4}, which includes line-of-sight absorbers at all distances, shows a large, 
blueshifted mean velocity field.
}
\label{fig:stacks_fg}
\end{figure}

@@

\begin{figure*}[!th]
\begin{minipage}[!bp]{0.28\textwidth}
\includegraphics[width=\textwidth]{Figures/fig_monopolar.pdf}
%\includegrpahics[width=\textwidth]{f07.pdf}
\caption{A cartoon showing a unipolar quasar. The gas observed in low- to intermediate-ion 
absorption preferentially lies behind the quasar, and is shadowed from the ionizing radiation. 
}
\label{fig:monopolar}
\end{minipage}%
\hspace{0.3in}
\begin{minipage}[!bp]{0.67\textwidth}
\includegraphics[width=\textwidth]{Figures/fig_lighttravel.pdf}
%\includegraphics[width=\textwidth]{f08.pdf}
\caption{A cartoon showing the finite lifetime of quasar episodes as an explanation to the 
asymmetric absorption. The setup on the left shows that the foreground quasar has not been shining 
long enough for its ionizing radiation to reach the gas behind it, when the light from the 
background quasar reaches. The setup on the right shows the scenario after an amount of time 
comparable to the light travelling time across CGM scale. Gas in front of the foreground quasar 
has been ionized, by the time the light from the background quasar reaches. 
}
\label{fig:lighttravel}
\end{minipage}
\end{figure*}

>>>>

\begin{figure}
\includegraphics[width=3.5in]{Figures/fig_stacks_and_fits.pdf}
%\includegraphics[width=3.5in]{f02.pdf}
\caption{Mean and median absorption centered at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 of the foreground quasars for all QPQ9 pairs. The composites are shown in thin, 
black. Overplotted on the composites are Gaussian fits, normalized to pseudo-continua far away 
from a velocity of $0\,{\rm km\,s^{-1}}$ relative to $z_{\rm fg}$, and are shown in thick, blue. 
For the doublets, a second Gaussian with a fixed mean separation and a tied standard deviation is 
included in the modeling. The absorptions frequently exhibit large velocity widths. The dashed 
lines mark the centroids, which show small positive velocity offsets from $z_{\rm fg}$. The 
$1\sigma$ modeling error for the centroid and dispersion of the \ion{C}{2} mean stack are 
$28\,{\rm km\,s^{-1}}$ and $29\,{\rm km\,s^{-1}}$ respectively, which is also the typical modeling 
errors of the other stacks.
}
\label{fig:stacks_and_fits}
\end{figure}

@@

In Figure~\ref{fig:stacks_and_fits}, we present mean and median stacks of \ion{C}{2}\,1334, 
\ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 absorption of the QPQ9 sample. We focus on the 
analysis results of the \ion{C}{2} mean stack. \ion{C}{4} and \ion{Mg}{2} are doublet transitions 
and it is more challenging to interpret their kinematics.
%To assess the average absorption of these transitions,
%we have combined the data with a novel technique.
%We have fitted the combined dataset using a cubic b-spline
%with knots at every 100\,kms.  We weighted each data
%point by the velocity width of the pixel to insure that 
%each sightline makes an equal contribution to the stack.
%These b-spline profiles are presented in the figure
%and a trivial integration gives the reported equivalent
%widths.
%[consider down-weighting systems with $<3\sigma$ EW]
%[Comnent on EW from 'standard' technique]
Two results are evident in Figure~\ref{fig:stacks_and_fits}: (i) the mean \ion{C}{2} stack 
exhibits excess absorption spanning a large velocity width; (ii) the mean absorption is likely 
skewed toward positive velocities. To model the absorption, we introduce a Gaussian profile while 
allowing a constant pseudo-continuum level to vary. We perform $\chi^2$ minimization with each 
channel given equal weight. From the best-fit to the data, we measure the centroid of the 
\ion{C}{2} stack to be $+208\,{\rm km\,s^{-1}}$ and the $1\sigma$ dispersion to be 
$318\,{\rm km\,s^{-1}}$. The centroid suggests an asymmetry that contradicts the standard 
expectation, while the dispersion suggests extreme kinematics. The median stack, on the other 
hand, shows weaker absorption, and the Gaussian model has a centroid with smaller offset and 
comparable dispersion.

We also create mean and median stack for the sub-sample with [\ion{O}{3}] redshifts and model the 
absorption with Gaussian best-fit. The \ion{C}{2} mean stack for this sub-sample has a centroid at 
$+249\,{\rm km\,s^{-1}}$, and a dispersion of $362\,{\rm km\,s^{-1}}$, consistent with the full 
sample.

@@

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_monte.pdf}
%\includegraphics[width=2.5in]{f05.pdf}
\caption{The same \ion{C}{2} mean stack shown in the first panel of 
Figure~\ref{fig:stacks_and_fits} is shown in thin, black. In thick, green, we overplot the 
Gaussian absorption model of the Monte Carlo simulations generated from a purely clustering 
argument. The model from clustering analysis is multiplied to the pseudo-continuum level of the 
stack of the observational data, shifted to the centroid of the stack of the data, and broadened 
by the mean redshift error in the data. The model is a good description of the data, with a 
dispersion within modeling error of the dispersion in the data. We overplot in dashed, yellow an 
absorption profile with $\sigma_v=573\,{\rm km\,s^{-1}}$, which is larger than the observed width 
by three times the standard deviation in the bootstrap analysis. Motions that produce a velocity 
width larger than this can be ruled out. We overplot in dashed, magenta an absorption profile with 
$\sigma_v=231\,{\rm km\,s^{-1}}$, which is smaller than the observed width by three times the 
modeling error. Unless gravitational and Hubble flows together with redshift error broadening 
produce a velocity width smaller than this, extra dynamical processes (e.g.\ outflows) will not be 
required to explain the observation. 
}
\label{fig:monte}
\end{figure}

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_contour.pdf}
%\includegraphics[width=2.5in]{f06.pdf}
\caption{
Probability distributions of the parameters $W_{\rm CII}$ and $M_{\rm halo}$. The plot shows the 
degeneracy between $W_{\rm CII}$ and $M_{\rm halo}$ in recovering the intrinsic width of the 
absorption profile. We mark contours for points that produce an absorption 
profile of width that is 1, 2, 3, and 4 times the modeling error away from the observed intrinsic 
width. The intrinsic velocity width corresponding to typical QPQ halo mass is marked with a plus 
sign, and is well contained within the 2$\sigma$ contour.}
\label{fig:contour}
\end{figure}

Under the assumption that the intrinsic dispersion and the redshift uncertainty add in 
quadrature to give the observed width, we solve for the intrinsic dispersion in the \ion{C}{2} 
mean stack. For the full QPQ9 sample, with the mean 
$\sigma_{{\rm error}(z)}^{\rm full}=182\,{\rm km\,s^{-1}}$, we recover 
$\sigma_{\rm intrinsic}^{\rm full}=261\,{\rm km\,s^{-1}}$. For the sub-sample with [\ion{O}{3}] 
redshifts, we recover $\sigma_{\rm intrinsic}^{\rm [OIII]}=356\,{\rm km\,s^{-1}}$.

\cite{Eftekharzadeh+15} measured the clustering of quasars in the range $2.2<z<2.8$ while 
\cite{RodriguezTorres+16} measured the clustering of quasars in the range $1.8<z<2.2$. They 
estimated that these quasars are hosted by dark matter halos with mass 
$M_{\rm halo}=10^{12.5}\,M_\odot$ and $M_{\rm halo}10^{12.6}\,M_\odot$ respectively. If dark 
matter halos hosting QPQ9 quasars have a characteristic mass $10^{12.6}\,M_\odot$ and follow an 
NFW profile \citep{NavarroFrenkWhite97} with concentration 
parameter $c=4$, at $z\approx2$ the maximum circular velocity is $345\,{\rm km\,s^{-1}}$. 
\cite{TormenBouchetWhite97} found that the maximum circular velocity is $\approx1.4$ times the 
maximum of the one-dimensional velocity dispersion $\sigma_{\rm 1D}$. Hence, the line-of-sight 
velocity dispersion typical of QPQ9 halos is $\sigma_{\rm 1D}=246\,{\rm km\,s^{-1}}$. We can 
investigate whether gravitational motions and Hubble flows are sufficient to reproduce the 
dispersion in the data, using Monte Carlo methods to simulate the absorption signals.

@@

In Figure~\ref{fig:monte}, we show a comparison of the observational data stack and the Gaussian 
absorption model of the Monte Carlo simulations. In the figure, the Gaussian absorption model is 
broadened by the mean redshift error by adding it in quadrature to the dispersion in the model. 
The resulting stack of mock spectra has a $1\sigma$ dispersion of $284\,{\rm km\,s^{-1}}$. This is 
about $1\sigma$ modeling error away from the intrinsic dispersion in the \ion{C}{2} mean stack for 
the full sample, and about 3 times the modeling error away from the intrinsic dispersion in the 
stack of the sub-sample with [\ion{O}{3}] redshifts. We also test for the 
sensitivity of this measured dispersion to the correlation functions adopted. The QPQ6 clustering 
analysis is performed on only the strongest absorber near $z_{\rm fg}$, and a low $R_\perp$ 
sightline may in fact intercept more than one optically thick absorber. We double the number of 
absorbers for each mock sightline, and find the measured dispersion only increases by several 
${\rm km\,s^{-1}}$. Thus, the hypothesis that the observed velocity width is only produced by a 
combination of gravitational motions and Hubble flows cannot be ruled out. Extra dynamical 
processes are not necessary to explain the large velocity fields.

From the bootstrap 
analysis of the observed dispersion, a width of $\sigma_v>573\,{\rm km\,s^{-1}}$ would be larger 
than the observed by three times the scatter (overplotted in Figure~\ref{fig:monte}). In other 
words, taking into account the redshift errors, motions in addition to gravitational and Hubble 
flows that will produce an intrinsic dispersion $\sigma_v>543\,{\rm km\,s^{-1}}$ can be ruled out.

On the contrary, 
$\sigma_{\rm 1D}<114\,{\rm km\,s^{-1}}$ together with Hubble velocities and 
broadening by redshift errors will result in a velocity width that is more than three times the 
modeling error away from the observed width (overplotted in Figure~\ref{fig:monte}). This implies 
that, unless the characteristic $M_{\rm halo}<10^{11.6}\,M_\odot$, 
additional dynamical processes are not required to explain the observed width.

In Figure~\ref{fig:contour}, We show the probability distributions of the degenerate parameters 
$W_{\rm CII}$ and $M_{\rm halo}$ in recovering the intrinsic width of the absorption profile. We 
require that the amplitude of the absorption is reproduced within 3 times its modeling error, and 
mark contours for points in ($W_{\rm CII}$, $M_{\rm halo}$)-space that produce an absorption 
profile of width within 1, 2, 3, and 4 times the modeling error of the observed intrinsic width. 
From the figure, a higher $W_{\rm CII}$ means the $M_{\rm halo}$ that will reproduce the observed 
width is higher. If there are no extra dynamical processes, the intrinsic velocity width 
corresponding to typical QPQ halo mass is well contained within the 2$\sigma$ contour.

@@

\begin{figure}
\includegraphics[width=3.4in]{Figures/fig_stacks_fg.pdf}
%\includegraphics[width=3.4in]{f07.pdf}
\caption{Mean stacks of the foreground quasar spectra at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 for the QPQ9 sample. For \ion{C}{2} and \ion{Mg}{2}, the mean absorption is 
weaker than that in the background stacks, and there is no evidence for an excess at $z_{\rm fg}$. 
The stack for \ion{C}{4}, which includes line-of-sight absorbers at all distances, shows a large, 
blueshifted mean velocity field.
}
\label{fig:stacks_fg}
\end{figure}

@@

\begin{figure*}[!th]
\begin{minipage}[!bp]{0.28\textwidth}
\includegraphics[width=\textwidth]{Figures/fig_monopolar.pdf}
%\includegrpahics[width=\textwidth]{f08.pdf}
\caption{A cartoon showing a unipolar quasar. The gas observed in low- to intermediate-ion 
absorption preferentially lies behind the quasar, and is shadowed from the ionizing radiation. 
}
\label{fig:monopolar}
\end{minipage}%
\hspace{0.3in}
\begin{minipage}[!bp]{0.67\textwidth}
\includegraphics[width=\textwidth]{Figures/fig_lighttravel.pdf}
%\includegraphics[width=\textwidth]{f09.pdf}
\caption{A cartoon showing the finite lifetime of quasar episodes as an explanation to the 
asymmetric absorption. The setup on the left shows that the foreground quasar has not been shining 
long enough for its ionizing radiation to reach the gas behind it, when the light from the 
background quasar reaches. The setup on the right shows the scenario after an amount of time 
comparable to the light travelling time across CGM scale. Gas in front of the foreground quasar 
has been ionized, by the time the light from the background quasar reaches. 
}
\label{fig:lighttravel}
\end{minipage}
\end{figure*}
